---
title: "2020"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 4
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, cache = F, fig.align = 'center')
```


```{r data}
library(tidyverse)

load("../results/state_results")
load("../results/state_simulations")
load("../results/p_biden")
load("../results/ec_sims")
load("../results/results_biden")
load("../results/results_trump")

state_ts <- read_csv("../results/state_ts.csv")
prior_results <- read_csv("../data/state_results_16.csv")
```

_Updated `r format(Sys.Date(), format="%B %dth, %Y")`._

# National

## current snapshot

```{r topline-ec}
biden <- round(quantile(ec_sims[, 2], c(0.05, 0.95)))
trump <- round(quantile(ec_sims[, 1], c(0.05, 0.95)))
```

Biden is projected to win between **`r biden[1]` - `r biden[2]`** electoral college votes.  Trump is projected to win between **`r trump[1]` - `r trump[2]`** electoral college votes.  In `r dim(ec_sims)[1]` simulations of the presidential election from the perspective of `r format(Sys.Date(), format="%B %dth, %Y")`, Biden wins **`r paste0(round(mean(ec_sims[, 2] > ec_sims[, 1])*100), "%")`** and Trump wins **`r paste0(round(mean(ec_sims[, 1] > ec_sims[, 2])*100), "%")`**.


```{r ecplot, results = T, fig.width = 10}
data_frame(
  electoral_votes = c(ec_sims[, 1], ec_sims[, 2]),
  candidate = c(rep("Trump", nrow(ec_sims)), rep("Biden", nrow(ec_sims)))
) %>% filter(electoral_votes != 268) %>%
  ggplot() +
  aes(x = electoral_votes, fill = candidate, y = ..density..) +
  geom_histogram(alpha = 0.7, bins = 538,
                 position = "identity") +
  scale_x_continuous(limits = c(0, 538),
                     breaks = c(seq(0, 538, 100), 270)) +
  scale_fill_manual(values = c("blue", "red")) +
  geom_vline(xintercept = 270, lty = 2, col = "black") +
  theme_minimal() +
  labs(x = NULL, y = NULL, fill = NULL,
       title = "distribution of electoral college votes") +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold", hjust = .5,
                                  color = "#007ba7"),
        plot.subtitle = element_text(size = 8),
        panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 11))

```

## over time

```{r over time}
ec_ts <- read_csv("../results/ec_ts.csv")

ec_ts %>%
  gather(metric, value, -date) %>%
  mutate(candidate = sapply(strsplit(metric, "_"), tail, 1),
         metric = sapply(strsplit(metric, "_"), head, 1)) %>%
  spread(metric, value) %>%
  ggplot() +
  aes(x = date, y = mean, ymin = lower, ymax = upper,
      col = candidate, fill = candidate) +
  geom_line() +
  geom_ribbon(col = NA, alpha = 0.3) +
  scale_x_date(limits = c(as.Date(min(ec_ts$date)), as.Date("2020-11-10")),
               date_breaks = "1 month",
               date_labels =  "%b, %Y") +
  scale_y_continuous(limits = c(0, 538),
                     breaks = c(0, 100, 200, 270, 300, 400, 500, 538)) +
  geom_vline(xintercept = as.Date("2020-11-03"), lty = 2, col = "darkgrey") +
  geom_hline(yintercept = 270, lty = 2, col = "darkgrey") +
  theme_minimal() +
  scale_color_manual(values = c("blue", "red")) +
  scale_fill_manual(values = c("blue", "red")) +
  labs(col = NULL, fill = NULL, y = NULL, x = NULL,
       title = "electoral college votes") +
  theme(panel.grid.minor = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(size = 15, face = "bold", color = "#007ba7", hjust = 0.5)) +
  guides(col = F)


```


# Who will win each state?


```{r stateresults, results = 'asis'}
p_biden %>%
  mutate(cat = case_when(p_biden >= 0.9 ~ "Likely Biden",
                         p_biden > 0.7 & p_biden < 0.9 ~ "Lean Biden",
                         p_biden <= 0.1 ~ "Likely Trump",
                         p_biden > 0.1 & p_biden < 0.3 ~ "Lean Trump", 
                         TRUE ~ "Uncertain")) %>%
  group_by(cat) %>%
  arrange(p_biden) %>%
  mutate(rank = row_number()) %>%
  ungroup() %>%
  mutate(rank = max(rank) - rank) %>%
  ggplot() +
  aes(x = reorder(cat, p_biden), y = rank, label = state, col = p_biden) +
  geom_text() +
  geom_text(alpha = 0.8, size = 4) +
  scale_x_discrete(position = "top") +
  theme_void() +
  theme(axis.text.x = element_text(size = 12, face = "bold", color = "#007ba7"),
        axis.text.y = element_blank(),
        plot.title = element_text(color = "#007ba7")) +
  labs(x = NULL, y = NULL) +
  guides(col = F) +
  scale_color_gradient(low = "red",  high = "blue")

```



```{r stateplot, results = TRUE, fig.height = 8, fig.align = 'center'}
lwd <- 1.1
cex <- 1.2
alpha <- 0.5

results_biden %>%
  ggplot() +
  aes(x = mean, y = reorder(state, mean),
      xmin = lower, xmax = upper) +
  geom_errorbarh(height = 0, col = "blue", lwd = lwd, alpha = alpha) +
  geom_point(data = results_trump, 
             aes(x = mean, y = state,
                 xmin = lower, xmax = upper),
             col = "red", cex = cex) +
  geom_errorbarh(data = results_trump,
                 aes(x = mean, xmin = lower, xmax = upper), 
                 height = 0, col = "red", lwd = lwd, alpha = alpha) +
  geom_point(col = "blue", cex = cex) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(0, 1, 0.25),
                     labels = paste0(seq(0, 100, 25), "%")) +
  labs(x = "vote share", y = NULL,
       title = "estimated vote share by state") +
  theme(plot.title = element_text(size = 15, face = "bold", hjust = 0.5, color = "#007ba7"),
        plot.subtitle = element_text(size = 7),
        axis.text.y = element_text(size = 9))
```



Note that uncertainties at this time are due to a) close races and/or b) few polls.  As more polls become available and election day nears, error bars should shrink and the range of reasonable potential outcomes should narrow.


```{r table}
DT::datatable(state_results, options = list(scrollX = TRUE, 
                                            scrollY = "200px",
                                            pageLength = 51))
```


```{r state-detail, results = TRUE}

plot_state_ts <- function(s) {
  p <- state_ts %>% 
    filter(state == s) %>% select(-state) %>%
    gather(metric, value, -date) %>%
    mutate(candidate = sapply(strsplit(metric, "_"), tail, 1),
           metric = sapply(strsplit(metric, "_"), head, 1)) %>%
    spread(metric, value) %>%
    ggplot() +
    aes(x = date, y = mean, ymin = lower, ymax = upper,
        col = candidate, fill = candidate) +
    geom_hline(yintercept = prior_results$dem[prior_results$state == s],
               lty = 9, col = "blue", alpha = 0.7) +
    geom_hline(yintercept = prior_results$rep[prior_results$state == s],
               lty = 9, col = "red", alpha = 0.7) +
    geom_line(aes(group = candidate), lwd = 1) +
    geom_ribbon(col = NA, alpha = 0.3) +
    scale_x_date(limits = c(as.Date(min(state_ts$date)), as.Date("2020-11-10")),
                 date_breaks = "1 month",
                 date_labels =  "%b, %Y") +
    scale_y_continuous(limits = c(0, 1),
                       breaks = seq(0, 1, 0.25),
                       labels = paste0(seq(0, 100, 25), "%")) +
    geom_vline(xintercept = as.Date("2020-11-03"), lty = 2, col = "darkgrey") +
    theme_minimal() +
    scale_color_manual(values = c("blue", "red")) +
    scale_fill_manual(values = c("blue", "red")) +
    labs(col = NULL, fill = NULL, y = NULL, x = NULL, title = NULL) +
    theme(panel.grid.minor = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(size = 15, face = "bold",
                                    color = "#007ba7", hjust = 0.5)) +
    guides(col = F, fill = F)
  print(p)
}

p_win <- function(s, c = "Biden") {
  oppo <- ifelse(c == "Biden", "Trump", "Biden")
  res <- state_simulations %>%
    filter(state == s) %>%
    summarise(pwin = mean(value[candidate == c] > value[candidate == oppo])) %>%
    pull(pwin) %>%
    round(3)
  res
}

```


```{r statests, results='asis'}
library(pander)
for (s in p_biden$state) {
  pwin <- p_win(s, "Biden")
  pwin <- case_when(pwin < 0.01 ~ "less than 1%",
                    pwin > 0.99 ~ "more than 99%", 
                    TRUE ~ paste0(round(pwin*100, 0), "%"))
   pander::pandoc.header(s, level = 2)
   pander::pandoc.p('\n')
   pander::pandoc.p(paste0('Biden is projected to receive between ',
                           round(100*state_ts$lower_biden[state_ts$state == s 
                           & state_ts$date == max(state_ts$date)], 0), '% and ',
                           round(100*state_ts$upper_biden[state_ts$state == s 
                           & state_ts$date == max(state_ts$date)], 0), 
                           '% of the vote in ', s, '.  ',
                           'Trump is projected to receive between ',
                           round(100*state_ts$lower_trump[state_ts$state == s 
                           & state_ts$date == max(state_ts$date)], 0), '% and ',
                           round(100*state_ts$upper_trump[state_ts$state == s 
                           & state_ts$date == max(state_ts$date)], 0), 
                           '%.  In ', dim(state_simulations)[1]/(51*3), 
                           ' simulations of the election in ', s, 
                           ', Biden wins ', pwin, '.'))
   plot_state_ts(s)
   pander::pandoc.p('\n')
}
```
